---
title: "P8106_hw2_xy2395"
author: "Jack Yan"
date: "3/19/2019"
output: html_document
---

```{r setup, include=T, message=F}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(mgcv)
set.seed(123123)
```

## Data Entry
```{r message=F}
concrete <- 
  read_csv("./data/concrete.csv") %>% 
  janitor::clean_names()
```

### (a) Scatter Plots
```{r}
# matrix of predictors
x <- model.matrix(compressive_strength~., concrete)[,-1]
# vector of response
y <- concrete$compressive_strength
```

```{r}
theme1 <- trellis.par.get()
theme1$plot.symbol$col <- rgb(.2, .4, .2, .5) 
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <- rgb(.8, .1, .1, 1) 
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2) 
trellis.par.set(theme1)

featurePlot(x, y, plot = "scatter", labels = c("","Y"),
            type = c("p"), 
            layout = c(4, 2)
            )
```

### (b) Polynomial Regression

```{r}
fit1 <- lm(compressive_strength ~ water, data = concrete)
fit2 <- lm(compressive_strength ~ poly(water,2), data = concrete) 
fit3 <- lm(compressive_strength ~ poly(water,3), data = concrete) 
fit4 <- lm(compressive_strength ~ poly(water,4), data = concrete) 

# Cross validation
trRows <- createDataPartition(y = concrete$compressive_strength,
                              times = 10,
                              p = .50,
                              list = FALSE)

ctrl1 <- trainControl(method = "cv", number = 10)
ctrl5 <- trainControl(method = "repeatedcv", repeats = 5, number = 10) 

lmFit1 <- train(compressive_strength ~ water, 
                data = concrete, 
                method = "lm", 
                trControl = ctrl1)
lmFit2 <- train(compressive_strength ~ poly(water, 2), 
                data = concrete, 
                method = "lm", 
                trControl = ctrl1)
lmFit3 <- train(compressive_strength ~ poly(water, 3), 
                data = concrete, 
                method = "lm", 
                trControl = ctrl1)
lmFit4 <- train(compressive_strength ~ poly(water, 4), 
                data = concrete, 
                method = "lm", 
                trControl = ctrl1)

resamp <- resamples(list(lm1 = lmFit1, lm2 = lmFit2, lm3 = lmFit3, lm4 = lmFit4))
summary(resamp)

anova(fit1, fit2, fit3, fit4)
```

The cross validation shows that when d = 4, the RMSE is the lowest, so we choose the df = 4 model. The ANOVA also indicates that the 4 degree model is the best. So in this case, cross validation and ANOVA both suggest that df = 4 is the best.

```{r}
p <- 
  ggplot(data = concrete, aes(x = water, y = compressive_strength)) + 
  geom_point(color = rgb(0.9, .4, .2, .5))

pred_df = tibble(
  water = concrete$water,
  pred1 = fit1$fit,
  pred2 = fit2$fit,
  pred3 = fit3$fit,
  pred4 = fit4$fit
)

p +
  geom_line(aes(x = water, y = pred1, color = '1'), data = pred_df) +
  geom_line(aes(x = water, y = pred2, color = '2'), data = pred_df) +
  geom_line(aes(x = water, y = pred3, color = '3'), data = pred_df) +
  geom_line(aes(x = water, y = pred4, color = '4'), data = pred_df) +
  scale_color_discrete(name = "DF")
```

The plot shows different polynomial fits to the data.

### (c) Smoothing Spline
```{r}
fit.ss <- smooth.spline(x = concrete$water, y = concrete$compressive_strength)
fit.ss$df
```

The degrees of freedom obtained by cross validation is `r fit.ss$df`.

```{r}
waterlims <- range(concrete$water)
water.grid <- seq(from = waterlims[1],to = waterlims[2], length.out = 2000)

fit.ss1 <- smooth.spline(x = concrete$water, y = concrete$compressive_strength, df=10)
fit.ss2 <- smooth.spline(x = concrete$water, y = concrete$compressive_strength, df=100)

pred.ss <- predict(fit.ss,
                   x = water.grid)
pred.ss1 <- predict(fit.ss1,
                   x = water.grid)
pred.ss2 <- predict(fit.ss2,
                   x = water.grid)
pred.ss.df <- data.frame(pred = pred.ss$y,
                         pred1 = pred.ss1$y,
                         pred2 = pred.ss2$y,
                         water = water.grid)

p + 
geom_line(aes(x = water, y = pred, color = '68.9'), data = pred.ss.df) + 
geom_line(aes(x = water, y = pred1, color = '10'), data = pred.ss.df) +
geom_line(aes(x = water, y = pred2, color = '100'), data = pred.ss.df) +
  scale_color_discrete(name = "DF") +
  theme_bw()
```

As the degrees of freedom increses, the curve becomes more flexible and wiggly.

### (d) GAM

```{r}
fit.gam = gam(compressive_strength ~ 
                s(cement) + s(blast_furnace_slag) + s(fly_ash) + s(water) + s(superplasticizer) +
                s(coarse_aggregate) + s(fine_aggregate) + s(age),  data = concrete)
summary(fit.gam)
```

All terms are significant, so we include all terms in the model.

```{r}
par(mfrow = c(2, 4))
plot(fit.gam)
```

The relationship between `cement` and the response is approximately linear. All the other terms have a nonlinear relation with the response.
